Methodology for Assessing the State of the Practice for Domain X

### 2020-11-06 StateOfPractice/Methodology/Methodology.tex
removed sustainability from section 2 of methodology
removed performance benchmarks from scope

\pmi{comment from meeting: remove performance benchmarks from scope of exercise because enough other people consider this. Add similar comment into introduction explaining (link to examples of performance considered elsewhere in similar projects??) why performance is not considered}

The scope of this document is to conduct a current state of practice analysis of software in a scientific computing domain for the purpose of understanding quality and assisting in future development. The methodology presented here is part of an overall assessment of the impact of model driven engineering and code generation on the sustainability of scientific computing software. The research proposal for this assessment, which contains a more detailed description of this state of practice exercise, can be found \href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ResearchProposal.pdf}{here}. The research questions that this exercise will answer are found in Section \ref{ResearchQuestions}. Section \ref{StepsAQDS} outlines the steps to be taken.

@@

\begin{enumerate}
\item What are the artifacts that current software packages consist of? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development of current software packages?
\item What are the pain points for developers working on SCS projects? What aspects of the existing processes, methodologies and tools do they consider could potentially be improved? What processes, methodologies and tools have they changed to reduce obstacles?
\item For research software developers, what specific actions are taken to address the following:
\begin{enumerate}
\item sustainability?\pmi{previous meeting discussion did consider removing this; I do not see any current questions to developers that best fit this quality; confirm remove or add new question to developers?}
\item usability \pmi{question 10 touches on this but will need to be reworded or broken up - ``What if any actions were taken to address usability issues?''}
\item traceability \pmi{question to developers 3}
\item modifiability \pmi{question to developers 9}
\item maintainability \pmi{question to developers 9}
\item correctness \pmi{question to developers 8}
\item understandability \pmi{question to developers 10 touches on this; need to reword question in terms of an action}
\item unambiguity \pmi{question to developers 11}
\item reproducibility \pmi{question to developers 12}
\item visibility/transparency \pmi{questions 1,3,4,5,6 build to this but a question would need to be modified to include something along the line of ``has this information been added to the documentation?''}
\end{enumerate} 
\item How does software designated as high quality by this methodology compare with top rated software by the community? \pmi{review}
\end{enumerate}

@@

\begin{enumerate}
\item Start with the state of practice research questions. (Section~\ref{ResearchQuestions}) \pmi{To be updated based on feedback}
\item Identify the domain. (Section~\ref{SecIdentifyDomain}) \pmi{In second review}
\item \emph{Domain Experts}: Create a top ten list of software packages in the domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal, research questions, measurement template, survey for short list projects, usability tests, performance benchmarks, maintainability experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list. (Section~\ref{SecInitialFilter}) \pmi{in final review}
\item \emph{Domain Experts}: Review domain software list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item \emph{Domain Experts}: Vet domain analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts}) 
\item Gather source code and documentation for each prospective software package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure}) \pmi{To be edited based on feedback}
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions to Developers.pdf}{Questions to Developers})
\item Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\item Performance benchmarks\pmi{note: this is still in consideration}
\item Modifiability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed}
\item Document answers for research questions.
\end{enumerate}

@@

\section{Rank Short List} \label{SecRankShortList}
An AHP pairwise comparison between the short list packages with respect to usability and modifiability survey results, as well as performance benchmarks, is conducted by the domain expert. The tool can be found here. The usability and modifiability survey can be found \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{here}. Performance benchmarks can be found here.
\pmi{are we adding a tool for this?}
\pmi{performance benchmarks will be domain specific, how to proceed with providing a resource/guidance?}

>>>>

\pmi{comment from meeting: remove performance benchmarks from scope of exercise because enough other people consider this. Add similar comment into introduction explaining (link to examples of performance considered elsewhere in similar projects??) why performance is not considered}
\pmi{do we have specific examples of other people considering performance benchmarks?}

The scope of this document is to conduct a current state of practice analysis of software in a scientific computing domain for the purpose of understanding quality and assisting in future development. We leave the assessment of performance of such software to other similar projects. The methodology presented here is part of an overall assessment of the impact of model driven engineering and code generation on the sustainability of scientific computing software. The research proposal for this assessment, which contains a more detailed description of this state of practice exercise, can be found \href{https://github.com/smiths/AIMSS/blob/master/OverallResearchProposal/ResearchProposal.pdf}{here}. The research questions that this exercise will answer are found in Section \ref{ResearchQuestions}. Section \ref{StepsAQDS} outlines the steps to be taken.

@@

\begin{enumerate}
\item What are the artifacts that current software packages consist of? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development of current software packages?
\item What are the pain points for developers working on SCS projects? What aspects of the existing processes, methodologies and tools do they consider could potentially be improved? What processes, methodologies and tools have they changed to reduce obstacles?
\item For research software developers, what specific actions are taken to address the following:
\begin{enumerate}
\item usability \pmi{question 10 touches on this but will need to be reworded or broken up - ``What if any actions were taken to address usability issues?''}
\item traceability \pmi{question to developers 3}
\item modifiability \pmi{question to developers 9}
\item maintainability \pmi{question to developers 9}
\item correctness \pmi{question to developers 8}
\item understandability \pmi{question to developers 10 touches on this; need to reword question in terms of an action}
\item unambiguity \pmi{question to developers 11}
\item reproducibility \pmi{question to developers 12}
\item visibility/transparency \pmi{questions 1,3,4,5,6 build to this but a question would need to be modified to include something along the line of ``has this information been added to the documentation?''}
\end{enumerate} 
\item How does software designated as high quality by this methodology compare with top rated software by the community? \pmi{review}
\end{enumerate}

@@

\begin{enumerate}
\item Start with the state of practice research questions. (Section~\ref{ResearchQuestions}) \pmi{To be updated based on feedback}
\item Identify the domain. (Section~\ref{SecIdentifyDomain}) \pmi{In second review}
\item \emph{Domain Experts}: Create a top ten list of software packages in the domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal, research questions, measurement template, survey for short list projects, usability tests, maintainability experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list. (Section~\ref{SecInitialFilter}) \pmi{in final review}
\item \emph{Domain Experts}: Review domain software list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item \emph{Domain Experts}: Vet domain analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts}) 
\item Gather source code and documentation for each prospective software package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure}) \pmi{To be edited based on feedback}
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions to Developers.pdf}{Questions to Developers})
\item Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\item Modifiability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed}
\item Document answers for research questions.
\end{enumerate}

@@

\section{Rank Short List} \label{SecRankShortList}

\pmi{comments from meeting: this part is to `enable conversation'; the ranking starts the conversation - `why do they like this software'; yes to tool question - spreadsheet; enable conversation regarding what's good about the specific softwares (also important highlights of domain software?)}

An AHP pairwise comparison between the short list packages with respect to usability and modifiability survey results is conducted by the domain expert. The tool can be found here. The usability and modifiability survey can be found \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{here}. 
\pmi{are we adding a tool for this?}
